---
title: "Recommendation System"
author: "Cyrus Cai"
date: "3/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
load("../rda/edx.rda")
```

# 1 Introduction
With the assumption that historical data contains inights for the future, we attempt to predict human behavior. The research is done in various areas, such as retail, movies, polls etc. 

## 1.1 The Chanllenge 
Today, I follow many processors and will build a recommendation system for movies. The model will predict what rating the exsiting user will give for a movie that he/she never watched before.

## 1.2 Dataset: MovieLens
I will use **MovieLens** as our dataset. The dataset is generated by GroupLens reseach lab. You can download the data through the [link](http://files.grouplens.org/datasets/movielens/ml-10m.zip).

```{r MovieLens head,echo=FALSE}
head(edx,5)
```

## 1.3 Goal
I will find a consice and yet accurate model. The project is aimed at an RMSE that is less than 0.865. The project is also aimed to produce an easy-to-follow and reproducible report. 

# 2 Analysis
Before modeling, I create two dataset: **edx** and **validation**. The **validation** is 10% of the full dataset, and is only used for final assessement of the model. It won't be used for anywhere else in the project. The **edx** data is going to be split into two subset, **train_set** and **test_set**.

## 2.1 The Simplest Model


### Performance
### Pros
### Cons

## 2.2 Model 2
### Performance
### Improvements
### Defects

## 2.3 Model 3
### Performance
### Improvements
### Defects


# 3 Results

## Model Summary

# 4 Conclusion

## Summary of the report
## Limitations
## Future Works

# 5 Reference

1 [R Markdown Cheat Sheet](https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)
2 [Data Science Courses from HarvardX](https://courses.edx.org/)
