---
title: "Recommendation System"
author: "Cyrus Cai"
date: "3/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
load("../rda/edx.rda")
load("../rda/train_set.rda")
load("../rda/test_set.rda")
```

# 1 Introduction
With the assumption that historical data contains inights for the future, we attempt to predict human behavior. The research is done in various areas, such as retail, movies, polls etc. 

## 1.1 The Chanllenge 
Today, I follow many processors and will build a recommendation system for movies. The model will predict what rating the exsiting user will give for a movie that he/she never watched before.

## 1.2 Dataset: MovieLens
I will use **MovieLens** as our dataset. The dataset is generated by GroupLens reseach lab. You can download the data through the [link](http://files.grouplens.org/datasets/movielens/ml-10m.zip).

```{r MovieLens head,echo=FALSE}
head(edx,5)
```

## 1.3 Goal
I will find a consice and yet accurate model. The project is aimed at an RMSE that is less than 0.865. The project is also aimed to produce an easy-to-follow and reproducible report. 

# 2 Analysis
Before modeling, I create two dataset: **edx** and **validation**. The **validation** is 10% of the full dataset, and is only used for final assessement of the model. It won't be used for anywhere else in the project. The **edx** data is going to be split into two subset, **train_set** and **test_set**.

## 2.1 The Simplest Model
The simplest model,  y_i=mu+e_i

```{r}
mu<-mean(train_set$rating)
```

### Performance
RMSE(mu,test_set$rating)

### Pros and Cons

The histogram shows that majority of the movies have residual more than 1

the scatter plot shows that the residual convert to 1 as the rating count increase.In other words, the more rating for a movie doesn't increase the prediction accuracy with this model

```{r}
temp_p <- test_set %>% mutate(residul=(rating-mu)^2) %>% group_by(movieId) %>% 
  summarise(r=sqrt(mean(residul)),count=n()) 

temp_p  %>% 
  ggplot(aes(r))+geom_histogram()+labs(x="Residual",y="Movie Count",title="Histogram Movie Count vs Residual")

 temp_p%>% 
  ggplot(aes(count,r))+geom_point() +labs(x="Rating Count by Movie",y="Residual",title="Scatter Plot Residual vs Rating Count ")

```
the histogram shows that majority of the users have residual more than 1

the scatter plot shows that the residual convert to 1 as the rating count increase.In other words, the more rating from a user doesn't increase the prediction accuracy with this model

```{r}
temp_p <- test_set %>% mutate(residul=(rating-mu)^2) %>% group_by(userId) %>% 
  summarise(r=sqrt(mean(residul)),count=n())

temp_p%>% 
  ggplot(aes(r))+geom_histogram()+labs(x="Residual",y="User Count",title="Histogram User Count vs Residual")

temp_p %>% 
  ggplot(aes(count,r))+geom_point() +labs(x="Rating Count per User",y="Residual",title="Scatter Plot Residual vs Rating Count")

```
the bar chat shows that all the genres have residual greater than 1

the scatter plot shows that the residual doesn't decrease as the rating count increase.In other words, the more rating from a user doesn't increase the prediction accuracy with this model

```{r}
temp_p <- test_set %>% mutate(residul=(rating-mu)^2) %>% 
separate_rows(genres, sep = "\\|") %>% 
group_by(genres) %>% 
  summarise(r=sqrt(mean(residul)),count=n()) 

temp_p%>% ggplot(aes(genres,r))+geom_bar(stat = "identity") +
  labs(x="Genres",y="Residual",title="Bar Chart Residual vs Genres")+
  geom_text(aes(label=round(r,3)),position=position_dodge(width=0.9), vjust=-0.25)

temp_p%>%  ggplot(aes(count,r))+geom_point() +
  labs(x="Rating Count per Genres",y="Residual",title="Scatter Plot Residual vs Rating Count")
```


## 2.2 Model 2
### Performance
### Improvements
### Defects

## 2.3 Model 3
### Performance
### Improvements
### Defects


# 3 Results

## Model Summary

# 4 Conclusion

## Summary of the report
## Limitations
## Future Works

# 5 Reference

1 [R Markdown Cheat Sheet](https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)
2 [Data Science Courses from HarvardX](https://courses.edx.org/)
